An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-06 EconGrowth-R-Functions.Rnw
Fixed covariance tables to be complete, even for U columns and rows.

##############################
# Returns an xtable of covariance information for the specified country.
##
covarianceTable <- function(countryAbbrev){
  data <- loadData(countryAbbrev)
  # Grab only the data we want.
  dataForCovar <- cbind(data$iGDP, data$iCapStk, data$iLabor, data$iQ, data$iX)
  names <- c("$y$", "$k$", "$l$", "$q$", "$x$")
  if (countryAbbrev %in% countryAbbrevsU){
    # Add U data only if we have it.
    dataForCovar <- cbind(dataForCovar, data$iU)
    names <- c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")
  }
  covarResults <- cor(dataForCovar)
  colnames(covarResults) <- names
  rownames(covarResults) <- names
  covarXtable <- xtable(covarResults, 
                      caption=paste(countryAbbrev, "covariance table."), 
                      label=paste("tab:Covariance_", countryAbbrev, sep=""))
  return(covarXtable)
}

>>>>

##############################
# Returns an xtable of covariance information for the specified country.
##
covarianceTable <- function(countryAbbrev){
  data <- loadData(countryAbbrev)
  if (countryAbbrev %in% countryAbbrevsU){
    # Calculate covariances among the typical variables.
    dataForCovar <- cbind(data$iGDP, data$iCapStk, data$iLabor, data$iQ, data$iX, data$iU)
    # Calculates the covariances among y, k. l, q, x, and u using all available data, namely,
    # 1980-2011 for covariances among y, k, l, q, and x for developed economies
    # 1980-2000 for covariances involving u for developed economies
    # 1991-2011 for covariances among y, k, l, q, and x for developing economies
    covarResults <- data.frame(cor(x=dataForCovar, use="pairwise.complete.obs"))
    names <- c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")
  } else {
    dataForCovar <- cbind(data$iGDP, data$iCapStk, data$iLabor, data$iQ, data$iX)
    covarResults <- cor(dataForCovar)
    names <- c("$y$", "$k$", "$l$", "$q$", "$x$")
  }
  colnames(covarResults) <- names
  rownames(covarResults) <- names
  # Now make an xtable so it displays nicely
  covarXtable <- xtable(covarResults, 
                      caption=paste(countryAbbrev, "covariance table."), 
                      label=paste("tab:Covariance_", countryAbbrev, sep=""))
  return(covarXtable)
}

### 2013-05-06 EconGrowth.Rnw
Fixed covariance tables to be complete, even for U columns and rows.

\item{Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command? Partially complete: need to work on the $u$ data only.}

@@

@@

Figure \ref{fig:Historical_Lattice_Graph} presents historical data, including indexed values for output ($y$), capital stock ($k$), labor ($l$), and the various energy quantifications ($q$, $x$, and $u$) for 1980--2011 (developed economies) and 1991--2011 (other economies). Factors of production track each other closely, and high covariances among these variables can be seen in Tables \ref{tab:Covariance_US} through \ref{tab:Covariance_ZM}.

% You can add fig.pos="H" to force the figure position to be this exact place.

<<Historical_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

@@

<<CD_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType=NA)
@

@@

<<CDPartialResidualsQ, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for thermal energy ($q$).">>=
createPartialResidualPlot(modelType="CD", energyType="Q", countryAbbrev=NA)
@

<<CDPartialResidualsX, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for exergy ($x$).">>=
createPartialResidualPlot(modelType="CD", energyType="X", countryAbbrev=NA)
@

<<CDPartialResidualsU, fig.width=threePanelGraphWidth, fig.height=threePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for useful work ($u$).">>=
createPartialResidualPlot(modelType="CD", energyType="U", countryAbbrev=NA)
@

@@

<<CDeGridLattice, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Grid search over $\\gamma$ for Cobb-Douglas model with thermal energy ($q$).">>=
cdeGridLattice(energyType="Q")
@

@@

<<CDq_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="Q")
@

@@

<<CDx_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="X")
@

@@

<<CDu_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="U")
@

@@

<<CD_GDP_Lattice_Graph, fig.width=cdParameterGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Cobb-Douglas results.">>=
createCDLatticeGraph()
@

@@

<<CES_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType=NA)
@

@@

<<CESPartialResidualsQ, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for thermal energy ($q$).">>=
createPartialResidualPlot(modelType="CES", energyType="Q", countryAbbrev=NA)
@

<<CESPartialResidualsX, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for exergy ($x$).">>=
createPartialResidualPlot(modelType="CES", energyType="X", countryAbbrev=NA)
@

<<CESPartialResidualsU, fig.width=threePanelGraphWidth, fig.height=threePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for useful work ($u$).">>=
createPartialResidualPlot(modelType="CES", energyType="U", countryAbbrev=NA)
@

@@

<<CESq_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="Q")
@

@@

<<CESx_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="X")
@

@@

<<CESu_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="U")
@

@@

<<CES_GDP_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES results.">>=
createCESLatticeGraph()
@

@@

<<CIvsParam_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="95\\% Confidence interval widths vs. parameter values for key energy parameters ($\\gamma$, $\\delta$, and $\\sigma$) of the Cobb-Douglas and CES models with energy. If an economy appears in the region above a dashed line, the 95\\% confidence interval for the parameter extends to or beyond $\\gamma \\le$ 0 (cannot say with 95\\% confidence that energy is important for the Cobb-Douglas model), $\\delta \\ge$ 1 (cannot say with 95\\% confidence that energy is important for the CES model), or $\\sigma$ = 1 (cannot say with 95\\% confidence whether ($kl$) and ($e$) are complements or substitutes in the CES model).">>=
CIvsParamLattice(countryAbbrevScaling=0.6)
@

@@

<<LINEXq_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="Q")
@

@@

<<LINEXx_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="X")
@

@@

<<LINEXu_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="U")
@

@@

<<LINEX_GDP_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="LINEX results.">>=
createLINEXLatticeGraph()
@

@@

@@

\begin{itemize}
  \item{The use of exergy, as opposed to thermal energy, in economic growth models is probably unnecessary. The exergy allocation factors are all close to unity (Table \ref{tab:Exergy_Allocation}). In the absence of very large changes in energy mix for an economy, indexed exergy ($x$) and thermal energy ($q$) are very similar. Figure \ref{fig:Historical_Lattice_Graph} shows that $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{Figure \ref{fig:CDeGridLattice} shows that the set of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that minimizes $SSE$ for the energy-augmented Cobb-Douglas production function always (with the exception of Saudi Arabia) has $\beta = 0$ or $\gamma = 0$. This result implies that it is unnecessary to include both labor ($l$) and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model (Figures \ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and CES model (Figures \ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) indicate that augmenting the Cobb-Douglas and CES production functions with energy is not beneficial.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters for both the energy-augmented Cobb-Douglas ($\gamma$) and CES ($\delta$) production functions have large confidence intervals such that we are unable to say (at the 95\% confidence level) that energy is an important factor for economic growth, vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} indicates that energy (in all quantifications, $q$, $x$, and $u$) is complementary to capital stock ($k$) and labor ($l$) for producing economic growth.}
  \item{Figures \ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} indicate that LINEX model fitting parameters $a_0$ and $c_t$ are significant (at the 95\% confidence level) for nearly all countries. The LINEX production function provides reasonable good fits to historical data. This evidence indicates that the LINEX production function (with energy ($e$) as its only factor of production) is a reasonable replacement for capital stock ($k$) and labor ($l$) in the Cobb-Douglas and CES production functions.}
  \item{The concept of useful work ($u$) has potential and should be explored beyond developed economies. Figure \ref{fig:Historical_Lattice_Graph} shows that $u$ grows at very different rates from $k$ or $l$ for the developed economies. The covariance values in Section \ref{sec:Covariance_Tables} show that $u$ has much lower covariances with $k$ and $l$ than $q$ or $x$ has with $k$ and $l$. Thus, $u$ may contain more information relative to $k$ and $l$ than $q$ or $x$ and may, therefore, be more useful in economic growth modeling than $q$ or $x$. ******* Check this assumption after we obtain covariance values for $u$. }
\end{itemize}

The evidence above leads to a re-conceptualization of the relationship among energy, capital stock, and labor: energy is \emph{embedded} in capital stock and labor, and, therefore, it is unwarranted and unnecessary to include \emph{both} energy and the capital stock/labor pair in production functions. The empirical evidence indicates that an energy-only production function (LINEX) also provides a reasonable model for economic growth.

************* The above section needs to be re-evaluated after the statistical work is tightened up. ****************

@@

************* Matt ended here. ****************

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\label{sec:Covariance_Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
\setcounter{figure}{0} % Restarts figure number counter at 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following tables show covariances among GDP ($y$) and factors of production ($k$, $l$, $q$, $x$, and $u$).

<<Print_Covariance_Tables, results='asis'>>=
#I wish I could say the following, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable(countryAbbrev="US"); printCovarTable(countryAbbrev="UK"); printCovarTable(countryAbbrev="JP")
printCovarTable(countryAbbrev="CN"); printCovarTable(countryAbbrev="ZA")
printCovarTable(countryAbbrev="SA"); printCovarTable(countryAbbrev="IR")
printCovarTable(countryAbbrev="TZ"); printCovarTable(countryAbbrev="ZM")
@

@@

<<SFk_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="K")
@

<<SFl_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="L")
@

<<SFq_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="Q")
@

<<SFx_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="X")
@

<<SFu_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="U")
@

@@

<<SF_GDP_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Single-factor results.">>=
createSFLatticeGraph()
@

>>>>

@@

\item{\st{Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command? Partially complete: need to work on the $u$ data only.}}

@@

Figure \ref{fig:Historical_Lattice_Graph} presents historical data, including indexed values for output ($y$), capital stock ($k$), labor ($l$), and the various energy quantifications ($q$, $x$, and $u$) for 1980--2011 (developed economies) and 1991--2011 (other economies). Factors of production track each other closely, and high covariances among these variables can be seen in Tables \ref{tab:Covariance_US} through \ref{tab:Covariance_ZM}. **************** Discuss the covariances in more detail after we finalize the covariances for U. *******************

<<Historical_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

@@

<<CD_Params_Graph, fig.pos="H", fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType=NA)
@

@@

<<CDPartialResidualsQ, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for thermal energy ($q$).">>=
createPartialResidualPlot(modelType="CD", energyType="Q", countryAbbrev=NA)
@

<<CDPartialResidualsX, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for exergy ($x$).">>=
createPartialResidualPlot(modelType="CD", energyType="X", countryAbbrev=NA)
@

<<CDPartialResidualsU, fig.pos="H", fig.width=threePanelGraphWidth, fig.height=threePanelGraphHeight, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for useful work ($u$).">>=
createPartialResidualPlot(modelType="CD", energyType="U", countryAbbrev=NA)
@

@@

<<CDeGridLattice, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Grid search over $\\gamma$ for Cobb-Douglas model with thermal energy ($q$).">>=
cdeGridLattice(energyType="Q")
@

@@

<<CDq_Params_Graph, fig.pos="H", fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="Q")
@

@@

<<CDx_Params_Graph, fig.pos="H", fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="X")
@

@@

<<CDu_Params_Graph, fig.pos="H", fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.cap="Cobb-Douglas (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph(energyType="U")
@

@@

<<CD_GDP_Lattice_Graph, fig.pos="H", fig.width=cdParameterGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Cobb-Douglas results.">>=
createCDLatticeGraph()
@

@@

<<CES_Params_Graph, fig.pos="H", fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType=NA)
@

@@

<<CESPartialResidualsQ, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for thermal energy ($q$).">>=
createPartialResidualPlot(modelType="CES", energyType="Q", countryAbbrev=NA)
@

<<CESPartialResidualsX, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for exergy ($x$).">>=
createPartialResidualPlot(modelType="CES", energyType="X", countryAbbrev=NA)
@

<<CESPartialResidualsU, fig.pos="H", fig.width=threePanelGraphWidth, fig.height=threePanelGraphHeight, fig.align="center", fig.cap="CES partial residuals plot for useful work ($u$).">>=
createPartialResidualPlot(modelType="CES", energyType="U", countryAbbrev=NA)
@

@@

<<CESq_Params_Graph, fig.pos="H", fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="Q")
@

@@

<<CESx_Params_Graph, fig.pos="H", fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="X")
@

@@

<<CESu_Params_Graph, fig.pos="H", fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.cap="CES (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCESParamsGraph(energyType="U")
@

@@

<<CES_GDP_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES results.">>=
createCESLatticeGraph()
@

@@

<<CIvsParam_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="95\\% Confidence interval widths vs. parameter values for key energy parameters ($\\gamma$, $\\delta$, and $\\sigma$) of the Cobb-Douglas and CES models with energy. If an economy appears in the region above a dashed line, the 95\\% confidence interval for the parameter extends to or beyond $\\gamma \\le$ 0 (cannot say with 95\\% confidence that energy is important for the Cobb-Douglas model), $\\delta \\ge$ 1 (cannot say with 95\\% confidence that energy is important for the CES model), or $\\sigma$ = 1 (cannot say with 95\\% confidence whether ($kl$) and ($e$) are complements or substitutes in the CES model).">>=
CIvsParamLattice(countryAbbrevScaling=0.6)
@

@@

<<LINEXq_Params_Graph, fig.pos="H", fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="Q")
@

@@

<<LINEXx_Params_Graph, fig.pos="H", fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="X")
@

@@

<<LINEXu_Params_Graph, fig.pos="H", fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.cap="LINEX (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph(energyType="U")
@

@@

<<LINEX_GDP_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="LINEX results.">>=
createLINEXLatticeGraph()
@

@@

****** This section needs to be re-evaluated after the statistical work is tightened up during Summer 2013. ******

@@

\begin{itemize}
  \item{The use of chemical exergy, as opposed to thermal energy, in economic growth models is probably unnecessary. Table \ref{tab:Exergy_Allocation} shows that exergy allocation factors are all close to unity. In the absence of very large changes in energy mix for an economy, indexed chemical exergy ($x$) and thermal energy ($q$) are very similar, and Figure \ref{fig:Historical_Lattice_Graph} shows that $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{Figure \ref{fig:CDeGridLattice} shows that the set of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that minimizes $SSE$ for the energy-augmented Cobb-Douglas production function always (with the exception of Saudi Arabia) has $\beta = 0$ or $\gamma = 0$. This result implies that it is unnecessary to include both labor ($l$) and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model (Figures \ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and CES model (Figures \ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) indicate that augmenting the Cobb-Douglas and CES production functions with energy is not beneficial.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters for both the energy-augmented Cobb-Douglas and CES production functions ($\gamma$ and $\delta$, respectively) have large confidence intervals such that we are unable to say (at the 95\% confidence level) that energy is an important factor for economic growth, vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} indicates that energy (in all quantifications: $q$, $x$, and $u$) is complementary to capital stock ($k$) and labor ($l$) for producing economic growth, according to the CES model.}
  \item{Figures \ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} indicate that LINEX model fitting parameters $a_0$ and $c_t$ are significant (at the 95\% confidence level) for nearly all countries. Figure \ref{fig:LINEX_GDP_Lattice_Graph} shows that the LINEX production function provides reasonably good fits to historical data. This evidence indicates that the LINEX production function (with energy ($e$) as its only factor of production) is a reasonable replacement the Cobb-Douglas and CES production functions which have capital stock ($k$) and labor ($l$) as factors of production.}
  \item{****** Discuss the AIC values for each of the model types. It could happen that the LINEX models have a higher AIC value than the CD or CES models. If so, that would mean that the LINEX model is actually preferred to the CD or CES models, primarily because it does a good-enough job of predicting economic output with fewer adjustable parameters (2) compared to CD (2), CDe (3), CES (4), CESe (6). Discuss with Randy Pruim during Summer 2013.}
  \item{The concept of useful work ($u$) has potential and should be explored beyond developed economies. Figure \ref{fig:Historical_Lattice_Graph} shows that $u$ grows at very different rates from $k$ or $l$ for the developed economies. The covariance values in Tables \ref{tab:Covariance_US}--\ref{tab:Covariance_ZM} show that $u$ has much lower covariances with $k$ and $l$ than $q$ or $x$ has with $k$ and $l$. Thus, $u$ may contain more information relative to $k$ and $l$ than $q$ or $x$ and may, therefore, be more useful in economic growth modeling than $q$ or $x$. ******* Check this assumption after we obtain covariance values for $u$.**********}
\end{itemize}

The evidence above leads to a re-conceptualization of the relationships among energy, capital stock, and labor: energy is \emph{embedded within} capital stock and labor, and, therefore, it is both unwarranted and unnecessary to include \emph{both} energy ($e$) and the capital stock/labor pair ($kl$) in a production function. The empirical evidence indicates that \emph{both} (a) capital stock/labor production functions (Cobb-Douglas and CES without energy) and (b) an energy-only production function (LINEX) provide a reasonable models for economic growth.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\label{sec:Covariance_Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
\setcounter{figure}{0} % Restarts figure number counter at 0 in this section.
\setcounter{equation}{0} % Restarts equation counter to 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following tables show covariances among GDP ($y$) and factors of production ($k$, $l$, $q$, $x$, and $u$). Covariances are calculated for the following years: 
\begin{itemize}
  \item{1980-2011 for covariances among $y$, $k$, $l$, $q$, and $x$ for developed economies;}
  \item{1980-2000 for covariances involving $u$ for developed economies; and}
  \item{1991-2011 for covariances among $y$, $k$, $l$, $q$, and $x$ for developing economies.}
\end{itemize}

<<Print_Covariance_Tables, results='asis', eval=TRUE>>=
#I wish I could say the following, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable(countryAbbrev="US"); printCovarTable(countryAbbrev="UK"); printCovarTable(countryAbbrev="JP")
printCovarTable(countryAbbrev="CN"); printCovarTable(countryAbbrev="ZA")
printCovarTable(countryAbbrev="SA"); printCovarTable(countryAbbrev="IR")
printCovarTable(countryAbbrev="TZ"); printCovarTable(countryAbbrev="ZM")
@

@@

<<SFk_Params_Graph, fig.pos="H", fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="K")
@

<<SFl_Params_Graph, fig.pos="H", fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="L")
@

<<SFq_Params_Graph, fig.pos="H", fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="Q")
@

<<SFx_Params_Graph, fig.pos="H", fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="X")
@

<<SFu_Params_Graph, fig.pos="H", fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph(factor="U")
@

@@

<<SF_GDP_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Single-factor results.">>=
createSFLatticeGraph()
@
